---
title: "Lab 03 - Nobel laureates"
author: "Tsion"
date: "02/03/2026"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

Here is a link to the [lab instructions](https://datascience4psych.github.io/DataScience4Psych/lab03.html).

### Exercise 1

```{r get-to-know-data}
glimpse(nobel)
```

1. The data has 1026 observations (rows) and 26 columns (variables). Each row represents a single Noble Laureate (specifying 26 different variables like firstname, lastname, year Noble won, date/city/country born,...)

### Exercise 2

```{r create-filtered-new-variable}
nobel_living <- nobel %>%     #create nobel_living from nobel then pipe
  filter(                     
    is.na(died_date),        #died_date is NA
    !is.na(country),         #country is available (not NA)
    gender != "org"          #gender not equal to organization 
  ) 
```

```{r confirm-new-variable}
nrow(nobel_living)
```


### Exercise 3

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...
